### 2023-04-04 response-to-audit-report.md

## Overview of Changes to Marlowe-Cardano Specification and Haskell (Plutus) Implementation

The audit report's main and high-severity concerns for the Plutus validators for Marlowe on Cardano fell into to four clusters:

1. Incorrect handling of negative deposits
    - "2.1.1 Negative deposits allow stealing funds"
2. Incomplete prevention of double satisfaction
    - "2.1.2 Contracts vulnerable to double satisfaction attacks"
3. Lax enforcement of invariants in Marlowe state
    - "2.1.7 Positive balances are not checked for the output state"
    - "2.1.8 Non-validated Marlowe states"
    - "2.1.9 Total balance of ending state uncomputed"
    - "2.1.10 Unchecked ending balance"
4. Difference semantics of association lists in Isabelle and Plutus
    - "2.1.12 Different insertion functions used in Isabelle and Haskell code"

Incorrect handling of negative deposits (item 1 above) was mitigated by the addition of a guard against this behavior in Marlowe's Plutus validator so that its behavior is consistent with Marlowe semantics. Property-based tests were also added to check the correctness of the mitigation.

Although the Marlowe validator had prevented double-satisfaction among multiple copies of the Marlowe validator script running in the same transaction, it did not prevent double-satisfaction in cases where the Marlowe validator ran alongside another Plutus validator in the same transaction (item 2 above). Double satisfaction is now prevented by enforcing that the Marlowe validator is the only Plutus script that runs during transactions that make payments to parties. This allows Marlowe contracts to coordinate with other Plutus scripts, but only under conditions where double satisfaction is impossible. Once again, property-based tests were added to check the correctness of this mitigation.

The Marlowe validator had made optimistic assumptions about its own correct operation and hence did not check certain invariants in order to save on Plutus execution costs (item 3 above). The Marlowe semantics validator has now been thoroughly hardened against corruption of initial or final state so that it ensures that the three state-invariants of positive accounts, non-duplication of state entries (accounts, choices, and bound values), and a total internal value matching the script UTxO's value. This hardening was completed without unduly increasing the size (in bytes) of the validator or its Plutus execution cost: such was verified with simulated and on-chain measurement of execution costs for a library of real-life Marlowe contracts. Notably, the enforcement of the final invariants protects against funds becoming permanently locked in a Marlowe contract due to a garbled state: even in cases where the implementation of Marlowe semantics or parts of the Plutus validator were flawed, at least some of the execution paths of a Marlowe contract remain viable, so funds could be released even under nominally impossible but catastrophic difficulties. The property-based test suite was significantly enhanced to check double-satisfaction situations.

@@

A new golden test for negative deposits was added in commit [bbb9b8fc](https://github.com/input-output-hk/marlowe-cardano/commit/bbb9b8fc21ad93b7bcdb3b60be0f2f9c921f6bec) and property-based tests for negative deposits were added in commits [bbb9b8fc](https://github.com/input-output-hk/marlowe-cardano/commit/bbb9b8fc21ad93b7bcdb3b60be0f2f9c921f6bec), [9d5443c5](https://github.com/input-output-hk/marlowe-cardano/commit/9d5443c5e9306eaf49ac5389df455e07555c9e04), [4b81ee96](https://github.com/input-output-hk/marlowe-cardano/commit/4b81ee96b1e7c74c4964cf25f6bb599c80d07a52), and [15446073](https://github.com/input-output-hk/marlowe-cardano/commit/15446073131c4098fed781cdcc701b89b77f6bd0).

### 2.1.2 Contracts vulnerable to double satisfaction attacks *(Severity: High)*

> **File `marlowe-cardano-specification.md`, `Constraint 15`**
> 
> No datum is required for outputs fulfilling payments to addresses generated by the evaluation of a Marlowe contract. This implies that these outputs are vulnerable to double satisfaction in transactions involving other contracts that pay to the same wallets. An example is discussed in […].
> 
> One way to strengthen the implementation is for the Marlowe validator to demand that outputs paid to addresses contain a datum that identifies the contract instance, like the `TxOutRef` of the validator UTxO being spent. Then cooperation with other contracts is possible without double satisfaction if the validators of the other contracts demand a different datum for their outputs.

The Marlowe semantics validator had prevented a limited form of double-satisfaction attack by preventing two Marlowe validators to execute in the same transaction. The audit report correctly highlights that such an attack could be feasible in cases where another Plutus validator (not the selfsame version of the Marlowe semantics validator) executes in a transactions where a payment to an address or to the Marlowe role-payout validator satisfies both the Marlowe semantics validator and the other validator.

@@

![#f03c15](https://placehold.co/15x15/f03c15/f03c15.png) TODO: @hrajchert will add text here, since it relates to the Isabelle specification.

### 2.1.7 Positive balances are not checked for the output state *(Severity: High)*

> **File `marlowe-cardano-specification.md`, `Constraint 13`** 
> 
> *Positive balances* are only checked for the input, not for the output Marlowe state. If the semantics are flawed, a transaction can produce an unspendable output that does not satisfy this constraint.
> 
> If such a transaction is accepted, no further evaluation will be possible since all subsequent transactions will be rejected due to the very same Constraint 13. This is an hypothetical attack vector, where a malicious actor could send a transaction to block a contract.

The Plutus code for the Marlowe semantics validator had been implemented with the optimistic assumption that semantics would not be flawed and that the additional execution cost of checking the output state was not warranted. However, as the audit report points out, such a flaw would result in the contract being forever blocked from further execution. Adding a check on the final output's validity now prevents a contract from ever reaching a fully blocked state. Even if the semantics or Plutus code were flawed, at least some execution paths in the contract would remain viable, so that the contract could eventually terminate and would not permanently lock funds.

@@

### 2.1.8 Non-validated Marlowe states *(Severity: High)*

> **File `marlowe-cardano-specification.md`, `Missing constraint`** 
> 
> The validator is not specified to check that the Marlowe states in the input and output datums are valid. This condition is necessary for the lemmas about the Marlowe semantics to be applicable. The Marlowe state could become invalid if there is a flaw in the implementation of the semantics.
> 
> It also could be possible for the Marlowe state to be invalid if someone pays an output to the Marlowe validator with an invalid Marlowe state. Though this problem could be addressed with off-chain checks that prevent sending transactions that spend outputs with invalid Marlowe states. If off-chain checks are used, a note in the specification about how this is handled would be helpful.
> 
> An example showing betrayed user expectations is discussed in […].
> 
> For a valid Marlowe state, the association lists for bound values, accounts, and choices have keys sorted and without duplicates.

Prompted by this finding in the audit report, the Marlowe semantics validator has been altered to enforce all invariants (positive accounts; non-duplication of accounts, choices, and bound values; a total internal account balance that exactly matches the value in the script's UTxO) on both the initial and final states. This prevents execution of a Marlowe contract that was inadvertently or purposefully created by off-chain code with an invalid initial state. It also prevents transactions that corrupt the final state; such corruption could only happen if there were a flaw in the Plutus validator or Marlowe semantics implementation, in which case that particular execution-path of the contract would be blocked. The [Marlowe Best Practices Guide](https://github.com/input-output-hk/marlowe-cardano/blob/main/marlowe/best-practices.md) and [Marlowe Security Guide](https://github.com/input-output-hk/marlowe-cardano/blob/main/marlowe/security.md) warn of the importance of off-chain code creating valid initial state for Marlowe contracts.

@@

> **File `MoneyPreservation.thy`, various functions ``**
> 
> `moneyInRefundOneResult`, `moneyInApplyResult`, `moneyInApplyAllResult`, `moneyInTransactionOutput`, and `moneyInPlayTraceResult` have strange meanings when the result is an error. Arguably, on error there is no money to retrieve, so the return type should be `(Token \times int) option` instead.
> 
> Some lemmas rely on this behavior to have equalities hold even in cases of errors, but the cost is that the meaning is so surprising that the reader may be confused by it. It would be more reliable to have explicit and weaker lemmas that assert equalities only when there are no errors.

![#f03c15](https://placehold.co/15x15/f03c15/f03c15.png) TODO: @hrajchert will add text here, since it relates to the Isabelle specification.

### 2.1.12 Different insertion functions used in Isabelle and Haskell code *(Severity: High)*

> **File `Semantics.hs`, Several functions** 
> 
> Where `MList.insert` is used in the Isabelle semantics, [AssocMap.insert](https://github.com/input-output-hk/plutus/blob/v1.0.0/plutus-tx/src/PlutusTx/AssocMap.hs#L147-L148) is used in the Cardano implementation. However, the functions are not equivalent as demonstrated by the following examples: […]
> 
> This renders the Isabelle lemmas inapplicable for the Cardano integration. The lemmas need to demand some properties of an `insert` function without fully spelling it out, or the Cardano integration needs to use `MList.insert` instead of `AssocMap.insert`.
> 
> Similarly, functions `AssocMap.delete` and `MList.delete` differ in behavior when the input map is not sorted: […]
> 
> Functions `AssocMap.lookup` and `MList.lookup` also differ in behavior when the input map is not sorted: […]
> 
> The following usage places were found:
> -   Line 395, `evalValue` depends on `moneyInAccount` which depends on `AssocMap.lookup`.
> -   Line 413, `evalValue` depends on `AssocMap.lookup`.
> -   Line 428, `evalObservation` depends on `AssocMap.member`.
> -   Line 456, function `updateMoneyInAccount` relies on `AssocMap.delete` and `AssocMap.insert`.
> -   Line 482, function `reduceContractStep` relies on `AssocMap.insert`.
> -   Line 567, function `applyAction` relies on `AssocMap.insert`.

The porting of `MList` from Isabelle to Plutus would have increased the size and execution cost of the Marlowe semantics validator beyond the present limits of the Cardano ledger. Work is underway to modify the Isabelle semantics and proofs to use a data structure similar to Plutus's `AssocMap`. Instead of modifying the Marlowe semantics validator to address this finding, all usages of `AssocMap` were manually reviewed to verify that no vulnerability was introduced by its use (as compared to `MList`) *provided that the precondition of no duplicate entries in the `AssocMap` held*: the source code was annotated with informal reasoning documenting the safety of `AssocMap`'s use. That manual review was augmented by a comprehensive set of property-based tests to check the behavior of `MList` (under the precondition that it was ordered and contained no duplicates) against `AssocMap` (under the precondition that it contained not duplicates) for all functions used in Marlowe semantics or in the Marlowe validator.

@@

Commit [a2ff6aa3](https://github.com/input-output-hk/marlowe-cardano/commit/a2ff6aa334efb7b7955408d29cf0e2edf06bca08) annotates the Marlowe semantics validator with reasoning as to the safe use of `AssocMap`. Commits [af380a29](https://github.com/input-output-hk/marlowe-cardano/commit/af380a292c0369bcc77d8584d81ea80a31326b61), [e2277677](https://github.com/input-output-hk/marlowe-cardano/commit/e2277677987f359dcd8754275eccd5b7c9a880a7), [a95a616a](https://github.com/input-output-hk/marlowe-cardano/commit/a95a616aca59bc7159d3ffa92487f8f021820bff), [b65ccfec](https://github.com/input-output-hk/marlowe-cardano/commit/b65ccfec3708484ff52514671e0e227c13f084f3), and [9e068b6a](https://github.com/input-output-hk/marlowe-cardano/commit/9e068b6a60103f2a40604a7eaea57c5f29b59180) implement the property-based tests for the behavior of the Plutus `AssocMap` against the Isabelle `MList`.

### 2.1.13 Missing specification tests *(Severity: Medium)*

> **File `Spec/Marlowe/Semantics/Compute.hs`** 
> 
> There are no tests for the properties in Section 3 of `specification-v3-rc1.pdf`. Besides checking that there are no translation mistakes, these properties would also help contrasting the assumptions in the Isabelle and the Haskell sides, like the meaning of validity of an association list, which is focused in the previous issue.

Subsequent to the commencement of the audit, a "test oracle" was developed that checks implementations of Marlowe semantics (such as the Plutus validator) against a reference implementation generated directly from the Isabelle specification. That generated Haskell implementation is backed by Isabelle proofs of the correctness of Haskell code generation by Isabelle. The test oracle is now applied to the Marlowe semantics implementation in Plutus that the validator use. The test oracle uses the reference implementation to generate test cases to challenge the Plutus implementation and then it checks the Plutus result against its own result. This oracle provides sufficient coverage to address this audit-reports concern regarding the lack of property-based tests for Section 3 of the Marlowe specification.

@@

> **File `SingleInputTransactions.thy`, lemma `beforeApplyAllLoopIsUseless`, line *270***
> 
> This lemma seems to say that `reduceContractUntilQuiescent` has no effect when composed with `applyAllLoop`, because `applyAllLoop` evaluates `reduceContractUntilQuiescent`, and `reduceContractUntilQuiescent` is idempotent.
> 
> A more descriptive name for this lemma could be `reduceContractUntilQuiescent_hasNoEffect_before_applyAllLoop`

![#f03c15](https://placehold.co/15x15/f03c15/f03c15.png) TODO: @hrajchert will add text here, since it relates to the Isabelle specification.

### 2.3.16 Unused and undocumented lemmas *(Severity: Low)*

> **Several Isabelle files, several lemmas**
> 
> Some lemmas are never used, and they would need comments motivating their presence:
> 1.  In file `MoneyPreservation.thy`, line 257, lemma `transferMoneyBetweenAccounts_preserves_aux`.
> 2.  In file `QuiescentResult.thy`
>     1.  Line 5, lemma `reduceOne_onlyIfNonEmptyState`
>     2.  Line 153, lemma `reduceContractUntilQuiescent_ifDifferentReduced`
> 3.  In file `PositiveAccounts.thy`, line 66, lemma `positiveMoneyInAccountOrNoAccount_sublist_gtZero`. Furthermore, it is identical to `positiveMoneyInAccountOrNoAccount_gtZero_preservation`, but with an additional assumption `money > 0`.
> 4.  In file `ValidState.thy`
>     1.  Line 9, lemma `valid_state_valid_choices`
>     2.  Line 13, lemma `valid_state_valid_valueBounds`
> 5.  In file `SingleInputTransactions.thy`, line 1214, lemma `traceListToSingleInput_isSingleInput`. It is mentioned in a commented out line in `StaticAnalysis.thy`. Furthermore, the lemma can be expressed more concisely as `$$\llparenthesis \mathit{interval} = \mathit{inte}, \mathit{inputs} = \mathit{inp\_h}\ \#\ \mathit{inp\_t} \rrparenthesis\ \#\ t = \mathit{traceListToSingleInput}\ \mathit{t2} % \mathrel{% \mbox{\fontfamily{cmr}\fontencoding{OT1}\selectfont=}}% \joinrel\Rightarrow\mathit{inp\_t} = []$$`

@@

> **File `SingleInputTransactions.thy`, lemmas `applyAllLoop_independet_of_acc_error1, applyAllLoop_independet_of_acc_error2`, lines *977, 987***
> 
> It is unclear what `acc` refers to in these lemma names, as the lemmas are about the independence of warnings and payments, not accounts.

![#f03c15](https://placehold.co/15x15/f03c15/f03c15.png) TODO: @hrajchert will add text here, since it relates to the Isabelle specification.

### 2.3.27 Verbose lemma statement `playTraceAuxIterative_base_case` *(Severity: Low)*

> **File `SingleInputTransactions.thy`, lemma `playTraceAuxIterative_base_case`, line *1063***
> 
> The statement of this lemma is very verbose. A more natural (and slightly stronger) formulation could be `$$\begin{aligned} &\mathit{playTraceAux}\ \mathit{txOut}\ [\ \llparenthesis \mathit{interval} = \mathit{inte}, \mathit{inputs} = [h] \rrparenthesis, \llparenthesis \mathit{interval} = \mathit{inte}, \mathit{inputs} = t \rrparenthesis \ ] \\ =\ &\mathit{playTraceAux}\ \mathit{txOut}\ [\ \llparenthesis \mathit{interval} = \mathit{inte}, \mathit{inputs} = h\ \#\ t \rrparenthesis \ ]\end{aligned}$$`

@@

> **File `TransactionBound.thy`, lemma `playTrace_only_accepts_maxTransactionsInitialState`, line *316***
> 
> This lemma seems like the main result of this file. Assuming it is an important result, we recommend writing it as a `theorem` rather than a `lemma`.

![#f03c15](https://placehold.co/15x15/f03c15/f03c15.png) TODO: @hrajchert will add text here, since it relates to the Isabelle specification.

### 2.3.29 Inconsistent style with assumptions *(Severity: Low)*

> **File `Timeout.thy`, lemmas `timedOutReduceContractUntilQuiescent_closes_contract, timedOutReduceContractStep_empties_accounts`, lines *201/204, 211/214***
> 
> These lemmas use the hypothesis `$\mathit{minTime}\ \mathit{sta} \leq \mathit{iniTime}$` and build a state `$\mathit{sta}\ \llparenthesis \mathit{minTime} := \mathit{iniTime} \rrparenthesis)$` while other lemmas simply say `$\mathit{minTime}\ \mathit{sta} = \mathit{iniTime}$`. Readability would be improved by presenting these lemmas in the same style as the others, or documenting the need for these distinct presentations via code comments.

@@

> **File `ValidState.thy`, lemma `playTrace_preserves_valid_state`, line *194***
> 
> This lemma seems like the main result of this file. Assuming it is an important result, we recommend writing it as a `theorem` instead.

![#f03c15](https://placehold.co/15x15/f03c15/f03c15.png) TODO: @hrajchert will add text here, since it relates to the Isabelle specification.

### 2.3.33 Unnecessary assumptions *(Severity: Low)*

### 2023-04-04 response-to-audit-report.md

> **File `PositiveAccounts.thy`, lemmas `addMoneyToAccountPositive_match, addMoneyToAccountPositive_noMatch`, lines *12, 23***
> 
> The assumptions `$$\forall x\ \mathit{tok}.\ \mathit{positiveMoneyInAccountOrNoAccount}\ x\ \mathit{tok}\ \mathit{accs}$$` in `addMoneyToAccountPositive_match` and `$$\mathit{money > 0}$$` in `addMoneyToAccountPositive_noMatch` are unnecessary.
> 
> **File `PositiveAccounts.thy`, lemma `reduceContractStep_gtZero_Refund`, line *93***
> 
> The lemma has an assumption that is mostly redundant.
> 
> ``` {.text escapeinside="||" mathescape="true"}
>   Reduced
>     ReduceNoWarning
>     (ReduceWithPayment (Payment party (Party party) tok2 money))
>     (state|$\llparenthesis$|accounts := newAccount|$\rrparenthesis$|) Close
>     =
>   Reduced wa eff newState newCont
> ```
> 
> A stronger lemma would be valid, which results from eliminating the assumption and changing the conclusion to `positiveMoneyInAccountOrNoAccount y tok3 newAccount`.
> 
> **File `QuiescentResult.thy`, lemma `reduceContractStepPayIsQuiescent`, line *17***
> 
> The assumption $\mathit{cont} = \texttt{Pay}\ \mathit{x21}\ \mathit{x22}\ \mathit{tok}\ \mathit{x23}\ \mathit{x24}$ is unnecessary.
> 
> **File `Timeout.thy`, lemma `timedOutReduce_only_quiescent_in_close_When`, line *43***
> 
> The assumption $\mathit{minTime}\ \mathit{sta} \leq \mathit{iniTime}$ is unnecessary.
> 
> **File `Timeout.thy`, lemma `timedOutReduce_only_quiescent_in_close`, line *122***
> 
> The assumption\ $\mathit{minTime}\ \mathit{sta} \leq \mathit{iniTime}$ is unnecessary. However, removing it will require the later proof `timedOutReduceContractLoop_closes_contract` to be adjusted.
> 
> **File `Timeout.thy`, lemma `timedOutReduceContractLoop_closes_contract`, lines *170, 173***
> 
> The assumptions $\mathit{minTime}\ \mathit{sta} \leq \mathit{iniTime}$ and $\mathit{minTime} \mathit{sta} = \mathit{iniTime}$ are both present. The former is redundant.
> 
> **File `TimeRange.thy`, lemma `reduceStep_ifCaseLtCt_aux`, line *234***
> 
> The assumption\ $\mathit{env} = \llparenthesis \mathit{timeInterval} = (a, b) \rrparenthesis$ is unnecessary.
> 
> **File `ValidState.thy`, lemma `reductionStep_preserves_valid_state_Refund`, line *29***
> 
> The assumption `$$\begin{aligned} \mathit{newState} = \llparenthesis &\mathit{accounts} = \mathit{newAccounts}, \mathit{choices} = \mathit{newChoices}, \\ &\mathit{boundValues} = \mathit{newBoundValues}, \mathit{minTime} = \mathit{newMinTime} \rrparenthesis\end{aligned}$$` is unnecessary.

@@

> **File `Semantics.hs`, Function `playTraceAux`, line *710***
> 
> The function in the Isabelle code takes a `TransactionOutputRecord` while the Haskell version takes a `TransactionOutput`. This means `TransactionError` cannot be an input to `playTraceAux` in Isabelle, possibly invalidating proofs about its properties.

The `playTrace` and `playTraceAux` functions are not used by the Marlowe validators, so no mitigation was made for this audit-report finding.

### 2.6.3 Variable names differ from Isabelle code *(Severity: Low)*

> **File `Semantics.hs`, Several functions** 
> 
> Different variable names in Isabelle and Haskell make comparison harder. It is less of an issue when only one variable has been renamed in a function, but multiple variable renames require carefully mapping between the names to avoid confusion. For example, the code of `reduceContractStep` in line 482 (`Pay` case) is hard to compare.
> 
> Other name changes include:
> -   Line 456, function `updateMoneyInAccount` uses variable `money` where the Isabelle code uses `amount` and omits naming the last parameter.
> -   Line 473, function `giveMoneyToPay` uses variables `amount` and `accounts` instead of `money` and `accountsV` as in the Isabelle code.
> -   Line 541, function `reductionLoop` uses variable `con` instead of `ncontract`.
> -   Line 300, the data type `TransactionInput` corresponds to the type `Transaction` in the Isabelle code.
> -   Line 313, the data type `TransactionOutput` is isomorphic but not identical to the homonymous data type in the Isabelle code.
> -   Line 439, function `refundOne` uses a variable `balance` where the Isabelle code uses `money`.
> -   Line 463, function `addMoneyToAccount` uses variable `accounts` where the Isabelle code uses `accountsV`.

@@

These cosmetic recommendations of the audit report have not been implemented in the Marlowe semantics validator, as the discrepancies do not strongly impact comparison.

### 2.6.4 Naming of functions and variables *(Severity: Low)*

> **File `Several files`, Several functions** 
> 
> Several functions or variables could have more descriptive or precise names, for example:
> -   `Scripts.hs:193`: `validateBalances` could be called `allBalancesArePositive`.
> -   `Scripts.hs:206`: `validateInputs` could be called `allInputsAreAuthorized`.
> -   `Scripts.hs:324`: `checkScriptOutputAny` could be called `noOutputPaysToOwnAddress`, as it checks that *no* outputs pay to the script address.
> -   `Semantics.hs:439`: `refundOne` is named somewhat confusingly, and understanding the name requires the context of `reduceContractStep` where the function is called. Perhaps a better name would be `dropWhileNonPositiveAndUncons`.
> -   `Semantics.hs:597`: the binding `tailCase` should rather be named `tailCases`.

@@

The aforementioned bindings `validateBalances`, `validateInputs`, and `tailCase` have been renamed to `allBalancesArePositive`, `allInputsAreAuthorized`, and `tailCases` in commit [1c96edfe](https://github.com/input-output-hk/marlowe-cardano/commit/1c96edfe7e22ecc499ed75dd175490a4f2427bdc). For historical reasons, `refundOne` has not been renamed. The function `checkScriptOutputAny` is no longer present due to other mitigations made to the validator.

### 2.6.5 Unused functions *(Severity: Low)*

> **File `Several files`, Several functions** 
> 
> Several functions are unused and perhaps should be removed:
> -   `Semantics.hs:741`: `contractLifespanUpperBound` does not seem to be used anywhere, including tests.
> -   `Semantics.hs:680`: `isClose` is not used in the rest of the codebase (besides checking its behavior via testing). It should either be removed, or comments justifying its existence should be included.
> 
> In addition to that, the functions `validateBalances` and `totalBalance` (defined at `Semantics.hs:755` and `:762`) are only used in `Scripts.hs` and never reused, so they should probably be moved to `Scripts.hs`.

@@

Although the validator does not use these functions, off-chain code in other Haskell modules does. For the time being, the `contractLifespanUpperBound` and `isClose` have been retained in `Language.Marlowe.Core.V1.Semantics`. They will be relocated in the future if validator and non-validator code is separated into different modules.

### 2.6.6 Comments *(Severity: Low)*

> **File `Semantics.hs`, Function `refundOne`, line *439***
> 
> The comment describing the function is overly concise, as it does not mention the function removing all non-positive accounts before the first positive one, and effectively `uncons`-ing the list.

Commit [47c9deb2](https://github.com/input-output-hk/marlowe-cardano/commit/47c9deb2047a240d78eaf62306278ecfbe624bea) elaborates the documentation for `refundOne`.

> **File `Semantics.hs`, Function `addMoneyToAccount`, line *461***
> 
> There is a typo in the comment: `accoun` is written instead of `account`.

@@

Commit [d1477610](https://github.com/input-output-hk/marlowe-cardano/commit/d14776101c22eac7fbc67893020f37cbc2a9e339) adds a comment that the use of `foldr` in the Plutus implementation is for efficiency.

### 2.6.10 Constraint implementations differ from description *(Severity: Low)*

> **File `marlowe-cardano-specification.md`, Section `Plutus Validator for Marlowe Semantics`** 
> 
> Some constraints mentioned in the specification are written in a different structure than the corresponding constraint in `Scripts.hs`. While such a discrepancy may be useful to minimize verbosity, a unified structure when possible would alleviate a side-by-side comparison. Examples of these differing structures include Constraint 6 and Constraint 14.

The discrepancies of implementation between the constraints in the specification and the Plutus code result from the emphasis on readability in the specification and of efficiency in the implementation. The need to minimize of validator size and execution cost in the implementation has sometimes resulted in less concise or less readable code. Note that the Plutus compiler can be quite sensitive to the precise expression of a constraint as Haskell code.

@@

### 2.6.11 Missing argument of `computeTransaction` *(Severity: Low)*

> **File `marlowe-cardano-specification.md`, Section `Relationship between Marlowe Validator and Semantics`** 
> 
> The specification mentions the output datum as the (fifth) argument for the `computeTransaction` function, while it is not an argument to it.

Commit [00a0c240](https://github.com/input-output-hk/marlowe-cardano/commit/00a0c2406665f06753aecf6dbb63933edb275436) moves the comment about the non-existent fifth argument to a separate sentence following the list of arguments.

@@

### 2.6.12 Missing `smallMarloweValidator` *(Severity: Low)*

> **File `marlowe-cardano-specification.md`, Various sections** 
> 
> The specification mentions `smallMarloweValidator` in a few places, but it is never mentioned in the source code.

@@

This incorrect reference is fixed in commit [3fac0d17](https://github.com/input-output-hk/marlowe-cardano/commit/3fac0d1749d107f13743756aa50c9723e776e391).

### 2.6.14 `MarloweParams` differs from the specification *(Severity: Low)*

> **File `Semantics.hs`, type `MarloweParams`, line *355***
> 
> The specification defines `MarloweParams` to contain just the payout validator hash, while the definition in the Haskell code contains just the roles currency symbol.

This misstatement is remedied in commit [8691f335](https://github.com/input-output-hk/marlowe-cardano/commit/8691f3351e55758cc750eee0800b1458f0a8c554).

@@

## 2.7 Haskell tests

### 2.7.1 More precise failure checks *(Severity: Medium)*

> **File `Spec/Marlowe/Plutus/Specification.hs`, `Various tests`** 
> 
> The tests use the functions `checkSemanticsTransaction` and `checkPayoutTransaction` to verify that various error conditions cause transactions to be rejected. These functions test that a transaction passes or fails, but when it fails, the functions do not consider the error cause. Checking the exact cause is necessary to ensure the transaction is rejected because of the intended reason and not because of some other error condition arising in a particular test case by coincidence.
> 
> The absence of this information makes it easier to accidentally produce a test that is not testing what is intended.

A future version of the Marlowe test suite for `checkSemanticsTransaction` and `checkPayoutTransaction` will instrument Plutus scripts for the Marlowe validators so that the precise cause of the test failure can be verified.

@@

### 2.7.2 Missing tests *(Severity: Medium)*

> **File `Spec/Marlowe/Semantics/Compute.hs`, ``** 
> 
> The following properties could additionally be tested for `computeTransaction`:
> -   payment subtracts from an account,
> -   deposit adds to an account,
> -   `INotify` input produces the expected continuation,
> -   `IChoice` input produces the expected continuation,
> -   the hash of a successfully applied merkleized input matches the hash of the merkleized case.
> 
> Some of these are tested in `Spec/Marlowe/Semantics/Functions.hs` already for auxiliary functions.

Commit [ee2e5222](https://github.com/input-output-hk/marlowe-cardano/commit/ee2e52220e25736af29a1f01144d5fafbe9eed76) adds a property-based test that choice and deposit continue as expected; commit [333eee29](https://github.com/input-output-hk/marlowe-cardano/commit/333eee293aa3afe95d0aa7f678c7e7b7bad660ea) tests that choice produces expected continuation; commit [477284c7](https://github.com/input-output-hk/marlowe-cardano/commit/477284c78095482eba63adc2e6090ea68553db9d) tests that that deposits add to an account; commit [d241669b](https://github.com/input-output-hk/marlowe-cardano/commit/d241669bfb7ec8dad1cc92a7ca6185d08229817e) tests that payout subtracts from an account; and commit [f33fc720](https://github.com/input-output-hk/marlowe-cardano/commit/f33fc7206522031c93253323f3b8ca083d8ca8c7) tests that merkleization continues as expected.

@@

> **File `Spec/Marlowe/Semantics/Functions.hs`, `Missing merkleization tests`** 
> 
> The properties in this module do not seem to be tested with merkleized contracts or inputs except for `checkGetContinuation`. More merkleization tests should be added.

Commits [2a936bf1](https://github.com/input-output-hk/marlowe-cardano/commit/2a936bf1bbb890cc811b966b46a10fd38314dc25) and [8f8183d7](https://github.com/input-output-hk/marlowe-cardano/commit/8f8183d7f14c6bdc3841a1999077ef01ebac4144) significantly increase the coverage of merkleization in validator tests by utilizing a much larger set of test contracts and adding a probability that almost any test will involve a merkleized contract.

> **File `Spec/Marlowe/Semantics/Compute.hs`, function `checkFixInterval`, lines *100-102***
> 
> The test `checkFixInterval` is never instantiated with an invalid interval that is in the past, meaning the function `fixInterval` is never tested for that case.

Commit [811f048b](https://github.com/input-output-hk/marlowe-cardano/commit/811f048ba93db10972641c0c235864db1d6014e7) adds a property-based test for an invalid interval in the past.

@@

## Appendix: Post-Audit Changes in Marlowe Validator

```bash
git diff 523f3d56f22bf992ddb0b0c8a52bb7a5a188f9e9 47c9deb2047a240d78eaf62306278ecfbe624bea marlowe/src/Language/Marlowe/{Core/V1/Semantics*,Scripts.hs}
```

```diff
diff --git a/marlowe/src/Language/Marlowe/Core/V1/Semantics.hs b/marlowe/src/Language/Marlowe/Core/V1/Semantics.hs
index e31b8c478..0d7e94a20 100644
--- a/marlowe/src/Language/Marlowe/Core/V1/Semantics.hs
+++ b/marlowe/src/Language/Marlowe/Core/V1/Semantics.hs
@@ -94,12 +94,12 @@ module Language.Marlowe.Core.V1.Semantics
   , TransactionError(..)
   , TransactionWarning(..)
     -- * Utility Functions
+  , allBalancesArePositive
   , contractLifespanUpperBound
   , isClose
   , notClose
   , paymentMoney
   , totalBalance
-  , validateBalances
     -- * Serialisation
   , currencySymbolFromJSON
   , currencySymbolToJSON
@@ -217,6 +217,26 @@ import Text.PrettyPrint.Leijen (comma, hang, lbrace, line, rbrace, space, text,
 data Payment = Payment AccountId Payee Token Integer
   deriving stock (Haskell.Eq, Haskell.Show)

@@

 -- | Marlowe transaction input.
@@ -319,6 +360,31 @@ data TransactionOutput =
     | Error TransactionError
   deriving stock (Haskell.Show)

+instance ToJSON TransactionOutput where
+  toJSON TransactionOutput{..} =
+    object
+      [
+        "warnings" .= txOutWarnings
+      , "payments" .= txOutPayments
+      , "state" .= txOutState
+      , "contract" .= txOutContract
+      ]
+  toJSON (Error err) = object ["transaction_error" .= err]
+
+instance FromJSON TransactionOutput where
+  parseJSON =
+    withObject "TransactionOutput"
+      $ \o ->
+        let
+          asTransactionOutput =
+            TransactionOutput
+              <$> o .: "warnings"
+              <*> o .: "payments"
+              <*> o .: "state"
+              <*> o .: "contract"
+          asError = Error <$> o .: "transaction_error"
+        in
+          asTransactionOutput <|> asError

 -- | Parse a validator hash from JSON.
 validatorHashFromJSON :: JSON.Value -> Parser ValidatorHash
@@ -404,12 +470,22 @@ evalValue env state value = let
                                    then 0
                                    else n `Builtins.quotientInteger` d
         ChoiceValue choiceId ->
+            -- SCP-5126: Given the precondition that `choices` contains no
+            -- duplicate entries, this lookup behaves identically to
+            -- Marlowe's Isabelle semantics given the precondition that
+            -- the initial state's `choices` in Isabelle was sorted and
+            -- did not contain duplicate entries.
             case Map.lookup choiceId (choices state) of
                 Just x  -> x
                 Nothing -> 0
         TimeIntervalStart    -> getPOSIXTime (fst (timeInterval env))
         TimeIntervalEnd      -> getPOSIXTime (snd (timeInterval env))
         UseValue valId       ->
+            -- SCP-5126: Given the precondition that `boundValues` contains
+            -- no duplicate entries, this lookup behaves identically to
+            -- Marlowe's Isabelle semantics given the precondition that
+            -- the initial state's `boundValues` in Isabelle was sorted
+            -- and did not contain duplicate entries.
             case Map.lookup valId (boundValues state) of
                 Just x  -> x
                 Nothing -> 0
@@ -425,6 +501,11 @@ evalObservation env state obs = let
         AndObs lhs rhs          -> evalObs lhs && evalObs rhs
         OrObs lhs rhs           -> evalObs lhs || evalObs rhs
         NotObs subObs           -> not (evalObs subObs)
+                                   -- SCP-5126: Given the precondition that `choices` contains no
+                                   -- duplicate entries, this membership test behaves identically
+                                   -- to Marlowe's Isabelle semantics given the precondition that
+                                   -- the initial state's `choices` in Isabelle was sorted and did
+                                   -- not contain duplicate entries.
         ChoseSomething choiceId -> choiceId `Map.member` choices state
         ValueGE lhs rhs         -> evalVal lhs >= evalVal rhs
         ValueGT lhs rhs         -> evalVal lhs > evalVal rhs
@@ -435,10 +516,16 @@ evalObservation env state obs = let
         FalseObs                -> False

### 2023-04-04 response-to-audit-report.md

--- | Pick the first account with money in it.
+-- | Pick the first account with money in it, discarding any accounts prior to that if they have a non-positive balance.
 refundOne :: Accounts -> Maybe ((Party, Token, Integer), Accounts)
 refundOne accounts = case Map.toList accounts of
     [] -> Nothing
+    -- SCP-5126: The return value of this function differs from
+    -- Isabelle semantics in that it returns the least-recently
+    -- added account-token combination rather than the first
+    -- lexicographically ordered one. Also, the sequence
+    -- `Map.fromList . tail . Map.toList` preserves the
+    -- invariants of order and non-duplication.
     ((accId, token), balance) : rest ->
         if balance > 0
         then Just ((accId, token, balance), Map.fromList rest)
@@ -447,18 +534,30 @@ refundOne accounts = case Map.toList accounts of
